Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-14 vertical-thermal-review.Rmd
Fig 3 variables in the table stated
Figure 3: Variables in the inset table need to be stated. I can guess what they are, but this should be explicit.
red, NIR, PAR language updates

Light, specifically the proportion of incident light and photosynthetically active radiation (PAR), decreases from the canopy top to the forest floor.
The profile shape of light is modified by leaf area density, leaf clumping, canopy height, and canopy structure across species and forest types [Fig. `r fig_NEON_vertical`a-d, Supporting Information Figure `r SIfig_NEON_all`, @koike_leaf_2001]. Canopy foliage absorb a large portion of PAR, and selectively filter incident light along the gradient, therefore altering the spectral characteristics of light received in the lower canopy layers. Along with decreasing PAR, this entails a decrease in red to far red ratio of light as it reaches the forest floor, where understories are enriched in near infra red radiation and absorb light in wavelengths of 700-1000nm [de Castro 2000; @poorterLeafOpticalProperties2000].
This gradient is more pronounced in dense canopies, including broad-leaf and mixed forests (e.g., SCBI, SERC and HARV, Fig. `r fig_NEON_vertical`d), than in open canopy forests, including many conifer forests [e.g., OSBS, PUUM, and WREF, Fig. `r fig_NEON_vertical`d, Supporting Information Figure `r SIfig_NEON_all`, @chazdonPhotosyntheticLightEnvironments1984; @aussenacInteractionsForestStands2000a; @smithSeasonalDroughtrelatedChanges2019; @tymenQuantifyingMicroenvironmentalVariation2017;  @parkerConsequencesEnvironmentalHeterogeneity2019; @bartemucciOverstoryInfluencesLight2006; @baldocchiSeasonalVariationEnergy1997]. 
Variability in the light environment decreases with height, with the shaded understory being highly influenced by sunflecks, or brief increases in solar radiation caused by small canopy gaps and wind-induced canopy movements [@waySunflecksTreesForests2012]. 
Likewise, the mid-canopy experiences a highly variable light environment due to the sun's passage across a dynamically structured canopy surface [@waySunflecksTreesForests2012].
The implication of this light gradient is the variation of leaf traits along the vertical gradient [Table 1], and that upper canopy leaves have higher potential photosynthetic rate (see section 4.2), but also greater propensity towards elevated $T_{leaf}$ (see sections 2.2, 4.2).

@@

![**Figure `r fig_leaf_T`. Theoretical expectations for variation in the difference between leaf and air temperatures, $T_{Leaf}-T_{air}$, in response to (a) shortwave radiation, (b) wind speed, (c) relative humidity, (d) leaf width, and (e) stomatal conductance.** Leaf temperatures were modeled using the  *tealeaves* R package of @muir_tealeaves_2019 parameterized to represent a broadleaf species (*Quercus rubra* L.) in a mesic temperate forest (Harvard Forest, Massachusetts, USA) under both humid and drought conditions, as detailed in Supplementary Information Methods `r appendix_tealeaves_methods`. In each scenario, the independent variable was allowed to vary while other parameters were held constant at the values given in the table of biophysical constants. ](leaf_energy_balance/fig3_leafenergy.png)

>>>>

Light, specifically the proportion of incident light and photosynthetically active radiation (PAR), decreases from the canopy top to the forest floor.
The profile shape of light is modified by leaf area density, leaf clumping, canopy height, and canopy structure across species and forest types [Fig. `r fig_NEON_vertical`a-d, Supporting Information Figure `r SIfig_NEON_all`, @koike_leaf_2001]. Canopy foliage absorb a large portion of PAR (400-700nm), and selectively filter incident light along the gradient, therefore altering the spectral characteristics of light received in the lower canopy layers. Along with decreasing PAR, there is a decrease in the red to far red ratio of light as it reaches the forest floor, where understories are enriched in near infrared radiation and absorb light in wavelengths of 700-1000nm [de Castro 2000; @poorterLeafOpticalProperties2000].
This gradient is more pronounced in dense canopies, including broad-leaf and mixed forests (e.g., SCBI, SERC and HARV, Fig. `r fig_NEON_vertical`d), than in open canopy forests, including many conifer forests [e.g., OSBS, PUUM, and WREF, Fig. `r fig_NEON_vertical`d, Supporting Information Figure `r SIfig_NEON_all`, @chazdonPhotosyntheticLightEnvironments1984; @aussenacInteractionsForestStands2000a; @smithSeasonalDroughtrelatedChanges2019; @tymenQuantifyingMicroenvironmentalVariation2017;  @parkerConsequencesEnvironmentalHeterogeneity2019; @bartemucciOverstoryInfluencesLight2006; @baldocchiSeasonalVariationEnergy1997]. 
Variability in the light environment decreases with height, with the shaded understory being highly influenced by sunflecks, or brief increases in solar radiation caused by small canopy gaps and wind-induced canopy movements [@waySunflecksTreesForests2012]. 
Likewise, the mid-canopy experiences a highly variable light environment due to the sun's passage across a dynamically structured canopy surface [@waySunflecksTreesForests2012].
The implication of this light gradient is the variation of leaf traits along the vertical gradient [Table 1], and that upper canopy leaves have higher potential photosynthetic rate (see section 4.2), but also greater propensity towards elevated $T_{leaf}$ (see sections 2.2, 4.2).

@@

![**Figure `r fig_leaf_T`. Theoretical expectations for variation in the difference between leaf and air temperatures, $T_{Leaf}-T_{air}$, in response to (a) shortwave radiation, (b) wind speed, (c) relative humidity, (d) leaf width, and (e) stomatal conductance.** Leaf temperatures were modeled using the  *tealeaves* R package of @muir_tealeaves_2019 parameterized to represent a broadleaf species (*Quercus rubra* L.) in a mesic temperate forest (Harvard Forest, Massachusetts, USA) under both humid and drought conditions, as detailed in Supplementary Information Methods `r appendix_tealeaves_methods`. In each scenario, the independent variable was allowed to vary while other parameters were held constant at the values given in the table of biophysical constants. Biohphysical constants in the the table include: shortwave radiation (swr), wind speed (ws), relative humidity (rh), leaf width (lw), stomatal conductance (gs), air temperature (tair). ](leaf_energy_balance/fig3_leafenergy.png)
